Notes on Semi-Supervised `Novelty Detection'

### 2011-05-03 semisupervised/notes.tex
more changes

Dora uses \emph{probabilistic graphical models} to represent all extracted knowledge from the world.
Those allow to easily model certain probabilistic properties such as conditional independence
and easy the process of several computational problems on the distributions.
Several tools exists for this type of graphical manipulation such as: \emph{Bayesian networks}, \emph{Markov Random Fields}
or \emph{Factor Graphs}.

Dora underlying graphical methods are based on \emph{Factor Graphs}\citep{kschischang2001factor}: bipartites graphs composed of variable and factor nodes representing the random variables and several factor functions between the variables that are used to decomposed the distribution.

Any probability distributuion can be represented by a \emph{factor graph} and we will denote the probability of a given set of variables outputs $x$ on a graph $G$ as $P_G(x)$.

@@

  \path [line] (fAppearance) -- (appearance);
  \path [line] (fShape) -- (shape);
  \path [line] (fSize) -- (size);
\end{tikzpicture}
\caption{Example of a factor graph used for Dora.}
\end{figure}

@@

      \node [var] (appearance) {Appearance}; &
      \node [var] (shape) {Shape}; &
      \node [var] (size) {Size}; \\
    };
  \path [line] (fAppearance) -- (appearance);
  \path [line] (fShape) -- (shape);
  \path [line] (fSize) -- (size);
\end{tikzpicture}
\caption{Factor graph to model unconditional probability. It assumes independency between\dots}
\end{figure}

Using the described $G$ and $G'$ we evaluated our novelty detectors in the distribution used before and compared with an optimal detector.
As possible to see in \autoref{fig:dora-data-simple} our approaximation to $P(x)$ althouth very basic was able to improve a novelty detection based only on conditional probability.

In order to generate a simple model we decided to model the unconditional density probability function also with a graphical model $G'$ and so we approximate $P(x)$ as $P_{G'}(x)$.

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{dora-data/roc-simple.pdf}
\caption{ROC curve for thresholds learned on samples from samples in the dora distribution}
\label{fig:dora-data-simple}
\end{figure}

Evaluating $P(x|G)/P(x|G')$...

Method:
\begin{enumerate}
\item Train a graph model for unlabelled data without the labelled variables.
\item Add the labelled variables and train the factors with the labelled data.
\item Use $P_G(x)$ as novelty threshold.
\end{enumerate}

@@

\clearpage
\subsection{How Dora models the world?}
\label{sec:dora-models}

Currently dora uses a \emph{probabilistic graphical model} to model the world.
All variables that the system see are added and connected in between with the
\emph{probabilistic factors} that the system is aware of.
Those factors are learned from the labelled data that the system is given when
training.

The proposal here is to also use such an approach to model the unconditional
probability.
Under that a novelty detection threshold can be easily implemented
reusing the \emph{probabilistic graphical model} $G$ that dora uses as
representation of the world:
\begin{description}
\item[$P(x|know)$] is obtained evaluating $P(G)$.
\item[$P(x)$] can be obtained by creating $G'$ as an instance of $G$ but with the
variable node we want to perform novelty detection removed and the factors associated
with it replaced with factors learned from the unlabelled data.
\end{description}

We point to an additional simplification that if the original graphical model factors
depending on the variable to perform novelty detection were previously split on a
conditional and unconditional factors. Then $G'$ could be obtained by simply removing
all the factors associated with the variable to perform novelty detection on.

>>>>

Dora uses \emph{probabilistic graphical models} to represent all extracted knowledge from the world.
Those allow to easily model certain probabilistic properties such as conditional independence
and easy the process of several computational problems on the distributions.
Several tools exists for this type of graphical manipulation such as: \emph{Bayesian networks}, \emph{Markov Random Fields}
or \emph{Factor Graphs}.
Those tools are especially useful since they are generative models and allow to compute the probability of any given set of variables $x$.

Dora underlying graphical methods are based on \emph{Chain Graphs} that can be translated into \emph{Factor Graphs}\citep{kschischang2001factor}: bipartite graphs composed of variable and factor nodes representing the random variables and several factor functions between the variables that are used to decomposed the distribution.
Any probability distribution can be represented by a \emph{factor graph} and we will denote the probability of a given set of variables outputs $x$ on a graph $G$ as $P_G(x)$.

@@

  \path [line] (fAppearance) -- (appearance);
  \path [line] (fShape) -- (shape);
  \path [line] (fSize) -- (size);
\end{tikzpicture}
\caption{Example of a factor graph $G$ used for Dora for modelling the labelled samples.}
\end{figure}

@@

      \node [var] (appearance) {Appearance}; &
      \node [var] (shape) {Shape}; &
      \node [var] (size) {Size}; \\
    };
  \path [line] (fAppearance) -- (appearance);
  \path [line] (fShape) -- (shape);
  \path [line] (fSize) -- (size);
\end{tikzpicture}
\caption{A sample factor graph to model the unconditional probability. It assumes independent properties. Factors are trained giving access to the whole set of unlabelled data.}
\end{figure}

Using the described $G$ and $G'$ we evaluated our novelty detectors in the distribution used before and compared with an optimal detector.
As possible to see in \autoref{fig:dora-data-simple} our approximation $P_{G'}(x)$ to $P(x)$ although very basic was able to improve a novelty detection.

\begin{figure}[h]
\centering
\includegraphics[width=0.75\textwidth]{dora-data/roc-simple.pdf}
\caption{ROC curve for thresholds learned on samples from samples in the dora distribution. As possible to see even a simple approach to model $P(x)$ was able to considerably improve our novelty detector.}
\label{fig:dora-data-simple}
\end{figure}

@@

@@
